---
title: "Working title: Effect of symmetry on quantity discrimination"
author: "Leslie Ng"
date: "2024-01-08"
output:
  html_document:
   toc: yes
   toc_float: yes
   collapsed: true 
   smooth_scroll: true
   depth: 3 
   highlight: tango # different theme for the appearance of the code
   theme: flatly
   code_folding: hide
self_contained: yes
mode: selfcontained
editor_options:
  chunk_output_type: console
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "../Output") })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# **Introduction**

In this study we tested if training bees using symmetrical/asymmetrical stimuli had an effect on their quantity discrimination. This document provides details on data processing, statistical analysis and figures of the original manuscript submitted for peer review.

The following code is written in the R programming language.

# **Install required packages**

```{r eval=FALSE}
install.packages("lme4") # linear mixed effect models
install.packages("ggplot2") # plot graphs
install.packages("tidyverse") # QOL features
install.packages("wesanderson") # color palette
install.packages("dplyr")  # data manipulation
install.packages("BFpack") # Bayes Factor for logistic regression
install.packages("ggThemeAssist") # Add-in to customize ggplot2 themes
install.packages("performance") # For assumption checking
install.packages("DHARMa") # For assumption checking]
install.packages("patchwork") # For combining figures together
```

# **Load packages**

```{r, message=FALSE, warning=FALSE}
library(lme4) # linear mixed effect models
library(ggplot2) # plot graphs
library(tidyverse) # QOL features
library(wesanderson) # color palette
library(dplyr)  # data manipulation
library(BFpack) # Bayes Factor for logistic regression
library(ggThemeAssist) # you may need to restart you session for this add-in to appear
library(performance) # For assumption checking
library(DHARMa) # For assumption checking
library(patchwork)
```

# **Data importing**

```{r, message=FALSE, results="hide"}

# Data importing -----------------------------------------------------------------------

test <- read.csv("../Data/sym_test.csv", header=TRUE)
training <- read.csv("../Data/sym_training.csv", header=TRUE)
summary(test) #view summary
summary(training)


# Data wrangling ------------------------------------------------------------------------

test$BEEID <- as.factor(test$BEEID) # treat as a categorical factor
test$TESTCONDITION <- as.factor(test$TESTCONDITION)

training$BEEID <- as.factor(training$BEEID) # treat as a categorical factor
training$SYM <- as.factor(training$SYM)

symmetry <-subset(test, TRAININGCONDITION == "SYM") #subset data
asymmetry <-subset(test, TRAININGCONDITION == "ASYM")

symsym <-subset(symmetry, TESTCONDITION == "SYM") #subset data
mean(symsym$CHOICE)

asymsym <-subset(asymmetry, TESTCONDITION == "SYM")
mean(asymsym$CHOICE)

symasym <-subset(symmetry, TESTCONDITION == "ASYM") #subset data
mean(symasym$CHOICE)

asymasym <-subset(asymmetry, TESTCONDITION == "ASYM")
mean(asymasym$CHOICE)

symtrans <-subset(symmetry, TESTCONDITION == "TRANSF") #subset data
mean(symtrans$CHOICE)

asymtrans <-subset(asymmetry, TESTCONDITION == "TRANSF")
mean(asymtrans$CHOICE)

symconflict <-subset(symmetry, TESTCONDITION == "CONFL") #subset data
mean(symconflict$CHOICE)

asymconflict <-subset(asymmetry, TESTCONDITION == "CONFL")
mean(asymconflict$CHOICE)

symconflict2 <-subset(symmetry, TESTCONDITION == "CONFLLARGE") #subset data
mean(symconflict2$CHOICE)

asymconflict2 <-subset(asymmetry, TESTCONDITION == "CONFLLARGE")
mean(asymconflict2$CHOICE)



STRAINING <-subset(training, SYM == "S") #subset data
ATRAINING <-subset(training, SYM == "A")
symmetrytest <-subset(test, TESTCONDITION == "SYM") #subset data
asymmetrytest <-subset(test, TESTCONDITION == "ASYM") #subset data
transtest <-subset(test, TESTCONDITION == "TRANSF") #subset data

```

# **Logistic regressions**

### *Test data*
The first part of the model name indicates the training condition, and the second part indicates the specific test e.g. symasymlm refers to the model on data for the asymmetrical test involving bees trained with symmetrical stimuli 
```{r message=FALSE, error=FALSE}
symsymlm <- glmer(CHOICE~ 1 + (1|BEEID), family = binomial, data = symsym)
summary(symsymlm)

asymsymlm <- glmer(CHOICE~ 1 + (1|BEEID), family = binomial, data = asymsym) 
summary(asymsymlm)

symasymlm <- glmer(CHOICE~ 1 + (1|BEEID), family = binomial, data = symasym)
summary(symasymlm)

asymasymlm <- glmer(CHOICE~ 1 + (1|BEEID), family = binomial, data = asymasym) 
summary(asymasymlm)

symtranslm <- glmer(CHOICE~ 1 + (1|BEEID), family = binomial, data = symtrans)
summary(symtranslm)

asymtranslm <- glmer(CHOICE~ 1 + (1|BEEID), family = binomial, data = asymtrans) 
summary(asymtranslm)

symconfllm <- glmer(CHOICE~ 1 + (1|BEEID), family = binomial, data = symconflict)
summary(symconfllm)

asymconfllm <- glmer(CHOICE~ 1 + (1|BEEID), family = binomial, data = asymconflict) 
summary(asymconfllm)

symtestcompare<- glmer(CHOICE~TRAININGCONDITION+(1|BEEID), family = binomial, data = symmetrytest) 
summary(symtestcompare)

asymtestcompare<- glmer(CHOICE~ TRAININGCONDITION+(1|BEEID), family = binomial, data = asymmetrytest) 
summary(asymtestcompare)

transfertestcompare<- glmer(CHOICE~ TRAININGCONDITION+(1|BEEID), family = binomial, data = transtest) 
summary(transfertestcompare)

```

### *Confidence intervals*

```{r results='hide', error=FALSE, message=FALSE, warning=FALSE}

confint(symsymlm) #confidence interval
symsymlmconfintlow<-(exp(0.1684812)/(1 + exp(0.1684812))) #reverse link function
symsymlmconfinthigh<-(exp(0.8137388)/(1 + exp(0.8137388)))

confint(asymsymlm)
asymsymlmconfintlow <-(exp(0.2992044)/(1 + exp(0.2992044)))
asymsymlmconfinthigh<-(exp(1.166141 )/(1 + exp(1.166141)))


confint(symasymlm)
symasymlmconfintlow <-(exp(0.1243281)/(1 + exp(0.1243281)))
symasymlmconfinthigh<-(exp(0.9110778 )/(1 + exp(0.9110778)))


confint(asymasymlm)
asymasymlmconfintlow <-(exp(0.536397)/(1 + exp(0.536397)))
asymasymlmconfinthigh<-(exp(1.456451 )/(1 + exp(1.456451)))

confint(symtranslm)
symtranslmconfintlow <-(exp(0.3327167)/(1 + exp(0.3327167)))
symtranslmconfinthigh<-(exp(1.188838 )/(1 + exp(1.188838)))


confint(asymtranslm)
asymtranslmconfintlow <-(exp(0.5579531)/(1 + exp(0.5579531)))
asymtranslmconfinthigh<-(exp(1.3551496 )/(1 + exp(1.3551496)))

confint(symconfllm)
symconfllmconfintlow <-(exp(-0.3037767)/(1 + exp(-0.3037767)))
symconfllmconfinthigh<-(exp(0.3553857)/(1 + exp(0.3553857)))

confint(asymconfllm)
asymconfllmconfintlow <-(exp(-0.1840665)/(1 + exp(-0.1840665)))
asymconfllmconfinthigh<-(exp(0.747081)/(1 + exp(0.747081)))


```

# **Generate figures**

### *Create dataframe for figures*

```{r message=FALSE, results='hide', error=FALSE}
symdframe <- data.frame(condition = rep(c("Symmetry","Asymmetry","Transfer","Conflict")), # Create dataframe
                legend = c("S","A","T","C"), 
                proportion = c(0.619,0.619,0.669,0.506),
                confintlow = c(symsymlmconfintlow,symasymlmconfintlow,symtranslmconfintlow,symconfllmconfintlow),
                confinthigh = c(symsymlmconfinthigh,symasymlmconfinthigh,symtranslmconfinthigh,symconfllmconfinthigh))



symmeanpoints <- symmetry %>%   # Calculate mean of each block for each bee
  group_by(BEEID, TESTCONDITION) %>% 
  summarize(prop = mean(CHOICE)) 
  colnames(symmeanpoints)[2] ="condition" # rename column to match dframe
  levels(symmeanpoints$condition) <- tolower(levels(symmeanpoints$condition)) # rename condition names to match 
  levels(symmeanpoints$condition) <- str_to_title(levels(symmeanpoints$condition)) # recapitalize condition names
  levels(symmeanpoints$condition) <- c('Asymmetry', 'Conflict', 'Conflict large', 'Symmetry', 'Transfer')
  symmeanpoints$condition <- factor(symmeanpoints$condition, levels=c('Symmetry', 'Asymmetry', 'Transfer', 'Conflict', 'Conflict large'))
  levels(symmeanpoints$condition)
  symmeanpoints <- subset(symmeanpoints, condition == "Symmetry" | condition == "Asymmetry" | condition == "Transfer" | condition == "Conflict")
symdframe$condition <- factor(symdframe$condition, levels = c('Symmetry', 'Asymmetry', 'Transfer', 'Conflict'))

                           
  
  asymdframe <- data.frame(condition = rep(c("Symmetry","Asymmetry","Transfer","Conflict")), # Create dataframe
                legend = c("S","A","T","C"), 
                proportion = c(0.663,0.713,0.713,0.563),
                confintlow = c(asymsymlmconfintlow,asymasymlmconfintlow,asymtranslmconfintlow,asymconfllmconfintlow),
                confinthigh = c(asymsymlmconfinthigh,asymasymlmconfinthigh,asymtranslmconfinthigh,asymconfllmconfinthigh))

asymmeanpoints <- asymmetry %>%   # Calculate mean of each block for each bee
  group_by(BEEID, TESTCONDITION) %>% 
  summarize(prop = mean(CHOICE)) %>%
  mutate(prop = ifelse(prop == 1, 0.95, prop))
  colnames(asymmeanpoints)[2] ="condition" # rename column to match dframe
  levels(asymmeanpoints$condition) <- tolower(levels(asymmeanpoints$condition)) # rename condition names to match 
  levels(asymmeanpoints$condition) <- str_to_title(levels(asymmeanpoints$condition)) # recapitalize condition names
  levels(asymmeanpoints$condition) <- c('Asymmetry', 'Conflict', 'Conflict large', 'Symmetry', 'Transfer')
  asymmeanpoints$condition <- factor(asymmeanpoints$condition, levels=c('Symmetry', 'Asymmetry', 'Transfer', 'Conflict', 'Conflict large'))
  levels(asymmeanpoints$condition)
  asymmeanpoints <- subset(asymmeanpoints, condition == "Symmetry" | condition == "Asymmetry" | condition == "Transfer" | condition == "Conflict")
asymdframe$condition <- factor(asymdframe$condition, levels = c('Symmetry', 'Asymmetry', 'Transfer', 'Conflict'))
  
```

# **Generate figure**

```{r}

symgraph <- ggplot(symdframe, aes(x = condition, y = proportion, fill = condition)) +
  geom_hline(yintercept = 0.5, linetype = "dashed", colour = "grey") +
  labs(x = "Condition", y = "Mean proportion of choices", fill = "Condition") +
  geom_bar(stat = "identity", position = "dodge", width = .7) + 
  geom_point(data = symmeanpoints, aes(x = condition, y = prop, colour = BEEID), 
             position = position_jitter(width = 0.15, height = 0.01), size = 2, colour = "#696969") +
  scale_fill_manual(values = wes_palette("GrandBudapest2", n = 4)) +
  geom_errorbar(aes(ymin = confintlow, ymax = confinthigh), width = .2) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1, suffix = NULL), limits = c(0, 1)) +
    scale_x_discrete(labels = c("Sym", "Asym", "Transf", "Confl")) +  # Change x-axis labels +
  theme(panel.background = element_blank(),
        axis.line = element_line(colour = "black"),
        aspect.ratio = 1 / 1,
        legend.position = "none",
        axis.title = element_text(size = 15),
        axis.text = element_text(size = 15),
        plot.margin = margin(0.3,0.3,0.3,0.6, "cm"),
        axis.title.y = element_text(vjust = 3),
        axis.title.x = element_text(margin = margin(t = 10)))
print(symgraph)



asymgraph <- ggplot(asymdframe, aes(x = condition, y = proportion, fill = condition))+
  geom_hline(yintercept = 0.5, linetype = "dashed", colour = "grey") +
  labs (x= "Condition", y = "Mean proportion of choices", fill="Condition") +
  geom_bar(stat = "identity", position = "dodge", width = .7) + 
  geom_point(data = asymmeanpoints, aes(x = condition, y = prop), position = position_jitter(width=0.15, height=0.01), size =2, colour = "#696969") +
  scale_fill_manual(values = wes_palette("GrandBudapest2", n = 4)) +
  geom_errorbar(aes(ymin = confintlow, ymax = confinthigh), width=.2) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1, suffix = NULL), limits = c(0,1)) +
  scale_x_discrete(labels = c("Sym", "Asym", "Transf", "Confl")) +  # Change x-axis labels +
  theme(panel.background = element_blank(),
        axis.line = element_line(colour = "black"),
        aspect.ratio = 1/1,
        legend.position = "none",
                axis.title = element_text(size = 15),
        axis.text = element_text(size = 15),
        axis.title.y = element_blank(),
        plot.margin = margin(0.3,0.3,0.3,0.3, "cm"),
        axis.title.x = element_text(margin = margin(t = 10)))

print(asymgraph)


```

### *Saving combined figure*

```{r}

figx <- symgraph + asymgraph
figx + plot_annotation(tag_levels = "A") & 
  theme(plot.tag.position = c(-0.1, 1),plot.tag = element_text(size = 16, face = "bold"))
ggsave("figx.pdf", device = pdf, height= 4.1, width=8.5, dpi = 300, path = "../Output")
ggsave("figx.png", device = png, height= 4.1, width=8.5, dpi = 300, path = "../Output")
```

# **Check Model fit for above models**

```{r warning=FALSE, error = FALSE, message=FALSE}

models <- list(
  symsym = symsymlm,
  asymsym = asymsymlm,
  symasym = symasymlm,
  asymasym = asymasymlm,
  symtrans = symtranslm,
  asymtrans = asymtranslm,
  symconfl = symconfllm,
  asymconfl = asymconfllm,
  symtestcompare = symtestcompare,
  asymtestcompare = asymtestcompare,
  transfertestcompare = transfertestcompare
)


# Check for overdispersion
for (model_name in names(models)) {
  cat(paste0("Overdispersion Check for ", model_name, "\n"))
  print(check_overdispersion(models[[model_name]]))
  cat("\n")
}

# Compute performance metrics for each model
performance_results <- lapply(models, model_performance)

# Capture and print markdown output for each model
for (model_name in names(performance_results)) {
  cat(paste0("### Performance Metrics for ", model_name, "\n\n"))
  
  # Capture the output in markdown format
  output <- capture.output(display(performance_results[[model_name]], format = "markdown", digits = 2, caption = NULL))
  
  # Print the captured output to the console
  cat(paste(output, collapse = "\n"))
  
  cat("\n\n")
}

# Loop for simulating residuals
for (model_name in names(models)) {
  cat(paste0("Simulated Residuals and Plot for ", model_name, "\n"))
  
  # Simulate residuals
  simulated_res <- simulateResiduals(fittedModel = models[[model_name]])
  
  # Plot simulated residuals
  plot(simulated_res)
  
  cat("\n\n")
}

```

# **Supplementary Training data**

```{r}
symtrlm <- glmer(CHOICE~ CNUM + (1|BEEID), family = binomial, data = STRAINING)
summary(symtrlm)
symtrlm2 <- glmer(CHOICE~ CNUM + (1|BEEID), family = binomial, data = STRAINING, nAGQ = 0) 
summary(symtrlm2)
STRAINING$CNUM2 <- scale(STRAINING$CNUM)
symtrlm3 <- glmer(CHOICE~ CNUM2 + (1|BEEID), family = binomial, data = STRAINING) 
summary(symtrlm3)

asymtrlm <- glmer(CHOICE~ CNUM + (1|BEEID), family = binomial, data = ATRAINING)
summary(asymtrlm)
asymtrlm2 <- glmer(CHOICE~ CNUM + (1|BEEID), family = binomial, data = ATRAINING, nAGQ = 0)
summary(asymtrlm2)
ATRAINING$CNUM2 <- scale(ATRAINING$CNUM)
asymtrlm3 <- glmer(CHOICE~ CNUM2 + (1|BEEID), family = binomial, data = ATRAINING)
summary(asymtrlm3)


# Compare AICc values

models_supp <- list(
  sym_training_nAGQ = symtrlm2,
  sym_training_scaled = symtrlm3,
  asym_training_nAGQ = asymtrlm2,
  asym_training_scaled = asymtrlm3)

# Compute performance metrics for each model
performance_results2 <- lapply(models_supp, model_performance)

# Capture and print markdown output for each model
for (model_name in names(performance_results2)) {
  cat(paste0("### Performance Metrics for ", model_name, "\n\n"))
  
  # Capture the output in markdown format
  output <- capture.output(display(performance_results2[[model_name]], format = "markdown", digits = 2, caption = NULL))
  
  # Print the captured output to the console
  cat(paste(output, collapse = "\n"))
  
  cat("\n\n")
}
```

# **Supplementary Test data**
Testing if a conflict test using large numbers (10 vs 10) has an effect on preference
```{r}
symconfl_large_lm <- glmer(CHOICE~ 1 + (1|BEEID), family = binomial, data = symconflict2)
summary(symconfl_large_lm)
summary(symconflict2)

asymconfl_large_lm <- glmer(CHOICE~ 1 + (1|BEEID), family = binomial, data = asymconflict2) 
summary(asymconfl_large_lm)
summary(asymconflict2)

confint(symconfl_large_lm)
symconfl_large_lmconfintlow <-(exp(-0.6812548)/(1 + exp(-0.6812548)))
symconfl_large_lmconfinthigh<-(exp(1.239418)/(1 + exp(1.239418)))

confint(asymconfl_large_lm)
asymconfl_large_lmconfintlow <-(exp(-0.5036269)/(1 + exp(-0.5036269)))
asymconfl_large_lmconfinthigh<-(exp(0.3303297)/(1 + exp(0.3303297)))

# Create dataframe for supplementary figure comparing performance in EE and SA

dframeconfl2 <- data.frame(TRAININGCONDITION = rep(c("Sym", "Asym")), #create dataframe
                        prop = c(0.54, 0.48),
                        confintlow = c(symconfl_large_lmconfintlow, asymconfl_large_lmconfintlow),
                        confinthigh = c(symconfl_large_lmconfinthigh, asymconfl_large_lmconfinthigh))

symasym_large <-subset(test, TESTCONDITION == "CONFLLARGE")


symasym_large$TRAININGCONDITION <- as.character(symasym_large$TRAININGCONDITION)

symasym_large$TRAININGCONDITION <- paste0(
  toupper(substr(symasym_large$TRAININGCONDITION, 1, 1)),
  tolower(substr(symasym_large$TRAININGCONDITION, 2, nchar(symasym_large$TRAININGCONDITION)))
)

largepoints <- symasym_large %>%   # calculate mean of each block for each bee
  group_by(BEEID, TRAININGCONDITION) %>% 
  summarize(prop = mean(CHOICE)) %>% 
  mutate(prop = ifelse(prop == 1, 0.95, prop)) %>%
  mutate(prop = ifelse(prop == 0, 0.05, prop))


conflict2graph <- ggplot()+
  labs (x= "Training condition", y = "Mean proportion of choices") +
  geom_bar(data = dframeconfl2, aes(x = TRAININGCONDITION, y = prop, fill = TRAININGCONDITION), stat = "identity", position = "dodge") + 
   geom_point(data = largepoints, aes(x = TRAININGCONDITION, y = prop), position = position_jitter(width=0.15, height=0.01), size =2, colour="#696969") +
  geom_errorbar(data= dframeconfl2, aes(x = TRAININGCONDITION, y = prop, ymin = confintlow, ymax = confinthigh), width=.2, position=position_dodge(.9)) +
  geom_hline(yintercept = 0.5, linetype = "dashed", colour = " black") +
  scale_fill_manual(values = wes_palette("GrandBudapest2", n = 2)) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1, suffix = NULL), limits = c(0, 1))  +
  theme(panel.background = element_blank(),
        axis.line = element_line(colour = "black"),
        aspect.ratio = 1/1, 
        axis.title.y = element_text(vjust = 5),
        axis.title.x = element_text(vjust = -3),
        plot.margin = margin(0.6,0.6,0.6,0.6, "cm")) + theme(axis.title = element_text(size = 15),
    axis.text.x = element_text(size = 15),
    axis.text.y = element_text(size = 15),
    axis.title.y = element_text(vjust = 5, size = 15)) + theme(legend.position = "none")
conflict2graph
ggsave("figSuppx.pdf", device = pdf, height= 5, width=5, dpi = 300, path = "../Output")
ggsave("figSuppx.png", device = png, height= 5, width=5, dpi = 300, path = "../Output")
```
